State of the Practice of Lattice Boltzmann Method Software Development

### 2021-06-25 StateOfPractice/Peter-Notes/Report/Report.tex
edits to Chapter 5 of LBM Report

\end{singlespace}
\newpage
\pagenumbering{arabic}
\section{Introduction - DONE SECTION 1st DRAFT}

@@

\section{Domain Analysis - DONE SECTION 1st DRAFT}\label{domainanalysis}

@@

\newpage
\section{Methodology - DONE SECTION 1st DRAFT}

@@

\newpage
\section{Results - DONE SECTION 1st DRAFT}\label{qualityresults}

@@

Table \ref{moduledata} shows file and line of code data of the software packages. Packages with a high reusability score do not have many LOC per text file, generally having a few hundred lines or less. This suggests that the source code of these packages is likely functionally modularized, and modules could be reused in other projects.

There was a strong focus on modularity when designing the waLBerla framework to enhance productivity, reusability, and maintainability \citep{bauer2021walberla}. Its software design has enabled waLBerla to be successfully applied in several projects as a basis for various extensions \citep{bauer2021walberla}.\\

\begin{onehalfspacing}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\hline
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\hline
			DL\_MESO (LBE) & 310 & 51 & 170223& 549\\
			ESPResSo &1309& 86 & 186700&143\\
			ESPResSo++ &5328& 66 & 969196&182\\
			HemeLB &1065& 48 & 95104&89\\
			laboetie &133& 1 & 48403&364\\		
			LatBo.jl &41& 0 & 42172&1029\\
			LB2D-Prime &82& 19 & 54755&668\\
			LB3D &99& 76 & 39766&402\\
			LB3D-Prime &23& 6 & 12944&563\\
			lbmpy&201&  28 & 46489  &231\\
			lettuce &62& 0 & 5529&89\\
			LIMBES &26& 1 & 4872&187\\
			Ludwig&859& 32 & 109811&128\\
			LUMA&312& 19 & 4370670&14000\\
			MechSys &324& 3 & 85543&264\\
			MP-LABS &307& 3 & 43124&140\\
			OpenLB &1104& 5 & 209034&189\\
			Palabos &1829& 67 & 547623&299\\
			pyLBM &258& 85& 32314&125\\
			Sailfish &632& 11 & 69398&110\\
			SunlightLB & 36& 1 & 7646& 212\\
			TCLB &535& 7 & 43226&81\\
			waLBerla & 2395 & 67 & 848146&353\\
			\hline
		\end{tabular}
		\captionof{table}{Module Data}\label{moduledata}
	\end{center}
\end{onehalfspacing}

@@

\subsection{Artifacts Present - DONE 1st DRAFT}\label{artifacts}

@@

\subsubsection{Less Common Artifacts}
The following artifacts were less commonly found in the 23 software packages that were tested. The top four AHP ranked packages have most of these artifacts.
At the time of data collection, only one (Palabos) of the four packages did not have a user manual or guide, but there was a broken link on the package website indicating that such an artifact might exist. This broken link was later fixed, but this is not reflected in our data. Despite the broken link, Palabos does have a detailed and informative website. Another one (Ludwig) of the top four packages does not appear to have publicly visible design documentation. A third package (OpenLB) from the list does not appear to use a version control system. It is possible that such a system is used by the package since its website notes package version numbers, but the artifacts do not explicitly state the use of such a system.

@@

\subsection{Tools Used - DONE 1st Draft}\label{tools}

@@

Most of the software packages use some sort of build automation tools, most commonly Make. They also all use various technical and domain specific libraries. Technical libraries include visualization (e.g. Matplotlib, ParaView, Pygame, VTK), data analysis (e.g. Anaconda, Torch), and message passing libraries (e.g. MPICH, Open MPI, PyZMQ). Domain specific libraries are scientific computing libraries (e.g. SciPy). Libraries that are explicitly stated in artifacts, and were not noted during our observations, may fall outside of these categories.

@@

Thirteen (57\%) of the 23 software packages use GitHub for managing the project, including nine (64\%) of the 14 alive packages, and four of the nine (44\%) dead packages. This could be indicative of a transition to such software development and version control tools for SCS. Typically there are several simultaneous development branches in these projects.

@@

\subsection{Pain Points - DONE 1st Draft}\label{painpoints}

@@

Developers were asked to comment on obstacles in their development process, obstacles encountered by users, and potential future obstacles. Several development process obstacles were identified.

@@

Difficulties with ensuring correctness were also noted by several developers. They indicated that tests are run on all new source code additions, testing both individual modules and the system to verify correctness. These tests compare the package output to known correct output using test cases. The developer of TCLB commented that the amount of testing data that is needed for some cases is a problem as free testing services do not offer capabilities to store and process such large amounts of data, and in-house testing solutions needed to be created to address this limitation. The solution for this has been to limit the size of the testing problems, and running tests in small batches of few iterations.

@@

A few potential future obstacles were noted. The developer of ESPResSo noted that their source code had been written with a specific application in mind and that due to this there was too much coupling between components of the source code. This results in technical debt, and would have an impact on future modifiability and reusability when trying to extend the software, and the code would need to be refactored.

@@

The developer comments have emphasized an importance on source code, while documentation seems to be of secondary importance. It must be stressed that improved management of documentation could benefit development and help eliminate some of the developer concerns that were raised. The use of automatic document generation tools that capture scientific and computing knowledge, and transform it into software artifacts, is advised. Drasil is an automatic document generation tool that is further discussed in Section \ref{conclusion}.

\subsection{Quality Recommendations - DONE 1st Draft}\label{qualityrecommentations}

@@

\begin{itemize}
	\item Include OS compatibility, including specific OS versions.
	\item Include complete installation instructions.
	\item Installation instructions should be written as if the user does not have any dependencies installed and is installing on a clean OS. This can be tested on a clean environment using a virtual machine.
	\item List all dependencies in the installation instructions, dependency versions, and how to install them.
	\item If possible, automate the installation of dependencies. Use tools such as Make. 
	\item Automate the installation process as much as possible. Use tools such as Make.
	\item The installation instruction should only be in one location.
	\item Include descriptive error messages for errors encountered during installation. This was done by several top ranked packages, including Ludwig and LUMA
	\item Include a way to validate the installation. This can be done using a custom script, or a test case that specifies expected output. This was done by several top ranked packages, including OpenLB, Ludwig, LUMA, and Palabos.
	\item Include instructions for uninstalling the software. These were included with pyLBM.
\end{itemize}

@@

\begin{itemize}
	\item Use a requirements specification document. A template 
	\item Make public (on package website or GitHub) the requirements specification document, or explicitly reference the domain theory that the software is designed from.
	\item Ensure the above information is easy to find. Consider adding it to the user manual. 
	\item Development teams should include both domain experts and experienced software developers.
	\item Use and make public (on package website or GitHub) detailed documentation. Consider using automatic document generation tools like Doxygen, Drasil, or Sphinx.
	\item Include detailed tutorials that include expected output, like the \href{https://www.walberla.net/doxygen/index.html}{waLBerla tutorials}.
	\item Use unit tests during development and make them public (on package website or GitHub). This was available for Ludwig.
	\item Modularize the source code, separate components, hide information  behind well defined interfaces.
	\item Use continuous integration tools and processes (Bamboo, Jenkins, and Travis CI) during development. This was done by top ranked packages ESPResSo and Ludwig.
\end{itemize}

@@

\begin{itemize}
	\item The system must provide descriptive error messages when it encounters unexpected input, including incorrect data types, empty input, and missing files or links. Eighteen of the software packages behaved in such reasonable fashion when tested with unexpected input.
\end{itemize}

@@

\begin{itemize}
	\item Include user hardware and software requirements documentation. Hardware requirements are rarely listed in the software packages that were assessed. Only those offering GPU processing (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB) mentioned any hardware requirements. On the other hand, some sort of software requirements were available for all packages, even if only some dependencies or a compatible operating system.
	\item Include a user tutorial that indicates the expected output. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, pyLBM.
	\item Provide a detailed user manual. It should identify elements of user interfaces, and identify all requirements to model a system. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item Provide documentation that details the background theory information, or provide a reference to such information.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and Palabos did this. The importance of specifying user characteristics is discussed in \citep{smith2007requirements}.
	\item Keep all documentation in one location.
	\item Do not duplicate artifact information.
	\item Maintain a user support model (Git, email, forum, FAQ)
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to shorted development time and improve usability.
\end{itemize}

@@

\begin{itemize}
	\item Keep artifacts updated.
	\item Ideally include most of the common and less common artifacts listed in Section \ref{artifacts}.
	\item Include version numbers and release notes for all major source code and artifact releases. The top five ranked software packages include release notes.
	\item Have a defined process for accepting contributions. The top five ranked software packages include information on how to contribute.
	\item Include documentation for making contributions to the project. 
	\item Use an issue tracker (Git, email, SourceForge, other) to manage bugs and changes. Issues should be regularly reviewed and closed. All but three (laboetie, lettuce, Sailfish) of the software packages that use Git have most of their issues closed.
	\item Use a version control system (GitHib, CVS). Four (ESPREsSo, Ludwig, LUMA, Palabos) of the top five ranked packages use a version control system.
	\item Source code needs to be well commented. Typically, more than 10 percent of LBM package source code is comments, as presented in Table \ref{gitrepodata}.
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Eliminate code duplication.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including maintainability.
	\item Consider the recommended points for addressing traceability.
\end{itemize}

@@

\begin{itemize}
	\item Adopt a coding standard and document it in artifacts, including some examples. A coding standard needs to be part of continuous integration. Only
	LUMA and HemeLB explicitly identify coding standards. 
	\item Use consistent, distinctive, and meaningful code identifiers. Nineteen of the 23 software packages use such code identifiers.
	\item Use symbolic constants. Twelve of the 23 software packages use such constants.
	\item Identify algorithms that are used in source code comments. Document them in the artifacts. Cite relevant external sources if needed. Domain algorithms are noted in the source code of 11 packages.
	\item Add meaningful comments. Indicate what is being done in each section of source code.
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Include user manuals that identify all technical and problem modeling requirements. Thirteen software packages have a user manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide. 
	\item Consider adding a FAQ section to the documentation. This helped resolve some usability issues for OpenLB.
\end{itemize}

@@

\begin{itemize}
	\item Update all relevant documentation when a change to the software is made.
	\item Use automatic document generation tools (Doxygen, Drasil, Sphinx) to limit the time spent on updating documentation.
	\item Provide a resource, like a developer's guide, that helps orient developers to the artifacts, source code, and system architecture so that they can better document changes.  
\end{itemize}

@@

\begin{itemize}
	\item Documentation must state all technical requirements and software dependencies. A missing dependency was a frequent cause of fault conditions during this assessment.
	\item Provide a detailed user manual. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item Provide a table of symbols in the developer guide that map to names used in the course code.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item The documentation should either explain the underlying CFD theories or provide a reference to appropriate resources.
	\item Consider asking users for feedback on the documentation. The developer of pyLBM noted that such feedback would be appreciated.
\end{itemize}

@@

answers last question of research questions:

How does software designated as high quality by this methodology compare	with top rated software by the community?

@@

This section will point to potential threats to the validity in the measurement section of data collection, for example from above:

>>>>

\end{singlespace}
\newpage
\pagenumbering{arabic}
\section{Introduction}

@@

\section{Domain Analysis}\label{domainanalysis}

@@

\newpage
\section{Methodology}

@@

\newpage
\section{Results}\label{qualityresults}

@@

Table \ref{moduledata} shows file and line of code data of the software packages. Packages with a high reusability score do not have many LOC per text file, generally having a few hundred lines or less. This suggests that the source code of these packages is likely functionally modularized, and modules could be reused in other projects.\\

	\begin{center}
		\begin{onehalfspacing}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\hline
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\hline
			DL\_MESO (LBE) & 310 & 51 & 170223& 549\\
			ESPResSo &1309& 86 & 186700&143\\
			ESPResSo++ &5328& 66 & 969196&182\\
			HemeLB &1065& 48 & 95104&89\\
			laboetie &133& 1 & 48403&364\\		
			LatBo.jl &41& 0 & 42172&1029\\
			LB2D-Prime &82& 19 & 54755&668\\
			LB3D &99& 76 & 39766&402\\
			LB3D-Prime &23& 6 & 12944&563\\
			lbmpy&201&  28 & 46489  &231\\
			lettuce &62& 0 & 5529&89\\
			LIMBES &26& 1 & 4872&187\\
			Ludwig&859& 32 & 109811&128\\
			LUMA&312& 19 & 4370670&14000\\
			MechSys &324& 3 & 85543&264\\
			MP-LABS &307& 3 & 43124&140\\
			OpenLB &1104& 5 & 209034&189\\
			Palabos &1829& 67 & 547623&299\\
			pyLBM &258& 85& 32314&125\\
			Sailfish &632& 11 & 69398&110\\
			SunlightLB & 36& 1 & 7646& 212\\
			TCLB &535& 7 & 43226&81\\
			waLBerla & 2395 & 67 & 848146&353\\
			\hline
		\end{tabular}
		\captionof{table}{Module Data}\label{moduledata}
		\end{onehalfspacing}
	\end{center}

There was a strong focus on modularity when designing the waLBerla framework to enhance productivity, reusability, and maintainability \citep{bauer2021walberla}. Its software design has enabled waLBerla to be successfully applied in several projects as a basis for various extensions \citep{bauer2021walberla}.

@@

\subsection{Artifacts Present}\label{artifacts}

@@

\subsubsection{Less Common Artifacts}\label{lesscommon}
The following artifacts were less commonly found in the 23 software packages that were tested. The top four AHP ranked packages have most of these artifacts.
At the time of data collection, only one (Palabos) of the four packages did not have a user manual or guide, but there was a broken link on the package website indicating that such an artifact might exist. This broken link was later fixed, but this is not reflected in our data. Despite the broken link, Palabos does have a detailed and informative website. Another one (Ludwig) of the top four packages does not appear to have publicly visible design documentation. A third package (OpenLB) from the list does not appear to use a version control system. It is possible that such a system is used by the package since its website notes package version numbers, but the artifacts do not explicitly state the use of such a system.

@@

\subsection{Tools Used}\label{tools}

@@

Most of the software packages use some sort of build automation tools, most commonly Make. They also all use various technical and domain specific libraries. Technical libraries include visualization (e.g. Matplotlib, ParaView, Pygame, VTK), data analysis (e.g. Anaconda, Torch), and message passing libraries (e.g. MPICH, Open MPI, PyZMQ). Domain specific libraries are scientific computing libraries (e.g. SciPy). Libraries that are not explicitly stated in artifacts, or were not noted during our observations, may fall outside of these categories.

@@

Thirteen (57\%) of the 23 software packages use GitHub for managing the project, including nine (64\%) of the 14 alive packages, and four of the nine (44\%) dead packages. Two projects (Palabos, WaLBerla) use GitLab. This could be indicative of a transition to such software development and version control tools for SCS. Typically there are several simultaneous development branches in these projects.

@@

\subsection{Pain Points}\label{painpoints}

@@

Developers were asked to comment on obstacles in their development process, obstacles encountered by users, and potential future obstacles.

@@

Difficulties with ensuring correctness were also noted by several developers. They indicated that tests are run on all new source code additions, testing both individual modules and the system to verify correctness. These tests compare the package output to known correct output using test cases. The developer of TCLB commented that the amount of testing data that is needed for some cases is a problem as free testing services do not offer capabilities to store and process such large amounts of data, and in-house testing solutions needed to be created to address this limitation. The solution for this has been to limit the size of the testing problems, and to run tests in small batches with few iterations.

@@

A few potential future obstacles were noted. The developer of ESPResSo noted that their source code had been written with a specific application in mind and that due to this there was too much coupling between components of the source code. This results in technical debt, having an impact on future modifiability and reusability when trying to extend the software, and the code would need to be refactored.

@@

The developer comments have emphasized an importance on source code, while documentation seems to be of secondary importance. It must be stressed that improving documentation could benefit development and help eliminate some of the developer concerns that were raised. The use of automatic document generation tools that capture scientific and computing knowledge, and transform it into software artifacts, is advised. Drasil is an automatic document generation tool that is further discussed in Section \ref{conclusion}.

\subsection{Quality Recommendations}\label{qualityrecommentations}

@@

\begin{itemize}
	\item Include OS compatibility, including specific OS versions.
	\item Provide complete installation instructions.
	\item Installation instructions should be written as if the user does not have any dependencies installed and is installing on a clean OS. This can be tested on a clean environment using a virtual machine.
	\item List all dependencies in the installation instructions, dependency versions, and how to install them.
	\item If possible, automate the installation of dependencies. Use tools such as Make. 
	\item Automate the installation process as much as possible. Use tools such as Make.
	\item The installation instruction should only be in one location.
	\item Include descriptive error messages for errors encountered during installation. This was done by several top ranked packages, including Ludwig and LUMA
	\item Provide a way to validate the installation. This can be done using a custom script, or a test case that specifies expected output. This was done by several top ranked packages, including OpenLB, Ludwig, LUMA, and Palabos.
	\item Provide instructions for uninstalling the software. These were included with pyLBM.
\end{itemize}

@@

\begin{itemize}
	\item Use a requirements specification document. A template is presented in \citep{smith2005new}. 
	\item Make public (on package website or GitHub) the requirements specification document, or explicitly reference the domain theory that the software is designed from.
	\item Ensure the above information is easy to find. Consider adding it to the user manual. 
	\item Development teams should include both domain experts and experienced software developers.
	\item Use and make public (on package website or GitHub) detailed documentation. Consider using automatic document generation tools like Doxygen, Drasil, or Sphinx.
	\item Provide detailed tutorials that include expected output, like the \href{https://www.walberla.net/doxygen/index.html}{waLBerla tutorials}.
	\item Use unit tests during development and make them public (on package website or GitHub). This was available for Ludwig.
	\item Modularize the source code, separate components, hide information  behind well defined interfaces.
	\item Use continuous integration tools (Bamboo, Jenkins, and Travis CI) and processes during development. This was done by top ranked packages ESPResSo and Ludwig.
\end{itemize}

@@

\begin{itemize}
	\item The system must provide descriptive error messages when it encounters unexpected input, including incorrect data types, empty input, and missing files or links. Eighteen of the software packages behaved reasonably when tested with unexpected input.
\end{itemize}

@@

\begin{itemize}
	\item Include user hardware and software requirements documentation. Hardware requirements are rarely listed in the software packages that were assessed. Only those offering GPU processing (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB) mentioned any hardware requirements. On the other hand, some sort of software requirements were available for all packages, even if only some dependencies or a compatible operating system were mentioned.
	\item Provide a user tutorial that indicates the expected output. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, pyLBM.
	\item Provide a detailed user manual. It should identify elements of user interfaces, and identify all requirements to model a system. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item Provide documentation that details the background theory information, or provide a reference to such information.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and Palabos did this. The importance of specifying user characteristics is discussed in \citep{smith2007requirements}.
	\item Keep all documentation in one location.
	\item Do not duplicate artifact information.
	\item Maintain a user support model (Git, email, forum, FAQ)
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to shorten development time and improve usability.
\end{itemize}

@@

\begin{itemize}
	\item Keep artifacts updated.
	\item Ideally include most of the common and less common artifacts listed in Section \ref{artifacts}.
	\item Include version numbers and release notes for all major source code and artifact releases. The top five ranked software packages include release notes.
	\item Have a defined process for accepting contributions, and make public documentation for making contributions to the project. The top five ranked software packages include information on how to contribute. 
	\item Use an issue tracker (Git, email, SourceForge, other) to manage bugs and changes. Issues should be regularly reviewed and closed. All but three (laboetie, lettuce, Sailfish) of the software packages that use Git have most of their issues closed.
	\item Use a version control system (GitHib, CVS). Four (ESPREsSo, Ludwig, LUMA, Palabos) of the top five ranked packages use a version control system.
	\item Source code needs to be well commented. Typically, more than 10 percent of LBM package source code is comments, as presented in Table \ref{gitrepodata}.
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Eliminate code duplication.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including maintainability.
	\item Consider the recommended points for addressing traceability.
\end{itemize}

@@

\begin{itemize}
	\item Adopt a coding standard and document it in artifacts, including some examples. A coding standard needs to be part of continuous integration. Only
	LUMA and HemeLB explicitly identify coding standards. 
	\item Use consistent, distinctive, and meaningful code identifiers. Nineteen of the 23 software packages use such code identifiers.
	\item Use symbolic constants. Twelve of the 23 software packages use such constants.
	\item Identify algorithms that are used in source code comments. Document them in the artifacts. Cite relevant external sources if needed. Domain algorithms are noted in the source code of 11 packages.
	\item Add meaningful comments. Indicate what is being done in each section of source code.
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Provide a user manual that identifies all technical and problem modeling requirements. Thirteen software packages have a user manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide. 
	\item Consider adding a FAQ section to the documentation. This helped resolve some usability issues for OpenLB.
\end{itemize}

@@

\begin{itemize}
	\item Update all relevant documentation when a change to the software is made.
	\item Use automatic document generation tools (Doxygen, Drasil, Sphinx) to limit the time spent on updating documentation.
	\item Provide a developer's guide to help orient developers to the artifacts, source code, and system architecture so that they can better document changes.  
\end{itemize}

@@

\begin{itemize}
	\item Documentation must state all technical requirements and software dependencies. A missing dependency was a frequent cause of fault conditions during this assessment.
	\item Provide a detailed user manual. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item Provide a table of symbols in the developer's guide that maps to names used in the course code.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item The documentation should either explain the underlying CFD theories or provide a reference to appropriate resources.
	\item Consider asking users for feedback on the documentation. The developer of pyLBM noted that such feedback would be appreciated.
\end{itemize}

@@

This comparison helps evaluate the methodology that was used, and assess the validity of the findings. Threats to the validity of this exercise are noted in Section \ref{threats}. The software package designations of this report are compared to package repository ranking metrics in Section \ref{repmetrics}, and to software recommended by a domain expert in Section \ref{domainexpertrecommend}.

\subsubsection{Repository Ranking Metrics}\label{repmetrics}

\begin{center}
	\begin{onehalfspacing}
		\begin{tabular}{ p{3.5cm}p{2.5cm}p{2.5cm}p{2.5cm} }
			\hline
			Name & Our Ranking & Repository Stars & Repository Watches \\
			\hline
			Ludwig&1& 19 & 6\\
			ESPResSo &2& 115 & 23\\
			Palabos &3& 14 & GitLab\\
			OpenLB &4& No Git & No Git\\
			LUMA&5& 26 & 10\\
			pyLBM &6& 57& 8\\
			DL\_MESO (LBE) & 7 & No Git & No Git \\
			waLBerla & 8 & 17 & GitLab\\
			Sailfish &9& 154 & 39\\
			laboetie &10& 4 & 5\\
			TCLB &11& 61 & 16\\
			MechSys &12& No Git & No Git\\
			lettuce &13& 8 & 2\\
			ESPResSo++ &14& 31 & 13\\
			MP-LABS &15& 7 & 2\\			
			SunlightLB & 16& No Git & No Git\\
			LB3D &17& No Git & No Git\\			
			LIMBES &18& No Git & No Git\\
			LB2D-Prime &19& No Git & No Git\\		
			HemeLB &20& 22 & 15\\
			lbmpy&21&  4 & 2  \\	
			LB3D-Prime &22& No Git & No Git\\	
			LatBo.jl &23& 4 & 5\\			
			\hline
		\end{tabular}
		\captionof{table}{Repository Ranking Metrics}\label{repometrics}
	\end{onehalfspacing}
\end{center}

Table \ref{repometrics} presents our LBM software package rankings along with the repository ranking metrics of each software package.

Eight packages do not use GitHub or GitLab, so they do not have a measure of repository stars. Looking at the repository stars of the other 15 packages, we can observe a slight pattern where packages that have been highly ranked by our assessment do have more stars than lower ranked packages. The second ranked package (ESPResSo) has the second most number of stars. The ninth ranked package (Sailfish) has the most number of stars. Packages designated as lower quality often do not use GitHub or GitLab, or have few stars, except for HemeLB which has 22 stars. Our assessment of this package might not be accurate.

The repository watches column contains even less data to compare since two of the packages (Palabos, waLBerla) use GitLab, which does not include this metric. The pattern that can be observed with this metric is very similar to that of the stars metric. The ninth ranked package (Sailfish) has the most number of watches. The second ranked package (ESPResSo) has the second most number of watches.

Besides missing data, another threat to the validity of this comparison is the varying ages of the repositories. Older packages have been able to accumulate stars and watches for longer than newer packages. The true quality of new packages may not be reflected in their stars and watches.

\subsubsection{Domain Expert Recommended Software}\label{domainexpertrecommend}

@@

This section examines potential threats to the validity of this state of the practice assessment. These can be categorized into methodology and data collection issues.

The measures listed in our measurement template may not be broad enough to accurately capture some qualities. For example, there are only two measures of surface robustness. The measurement of robustness could be expanded, as it currently only measures unexpected input. Other faults could be introduced, but could require a large investment of time to develop, and might not be a fair measure for all packages. Similarly, reusability is assessed along the number of code files and LOC per file. While this measure is indicative of modularity, it is possible that some packages have many files, with few LOC, but the files do not contain source code that is easily reusable. The files may be poorly formatted, or the source code may be vague and have ambiguous identifiers. Furthermore, the measurement of understandability relies on 10 random source code files. It is possible that the 10 files that were chosen to represent a software package may not be a good representation of the understandability of that package.

Regarding data collection, a risk to the validity of this assessment is missing or incorrect data. Some software package data may not have been measured due to technology issues like broken links. This issue arose with the measurement of Palabos, which had a broken link to its user manual, as noted in Section \ref{lesscommon}.

Some pertinent data may not have been specified in public artifacts, or may be obscure within an artifact or web-page. The use of unit testing and continuous integration was mentioned in the artifacts of only two (ESPResSo, Ludwig) packages. However, interviews suggested a more frequent use of both unit testing and continuous integration in the development processes than what was observed from the initial survey of the artifacts. For example, OpenLB, pyLBM, and TCLB use such methods during development despite this not being explicitly clear from an analysis of the material available online.

Furthermore, design documentation was measured to be a ``less common'' artifact in this assessment, but it is probable that such documentation is part of all LBM packages. After all, developing SCS is not a trivial endeavor. It is likely that many packages have such documentation but did not make it public, and due to this the measured data is not a true reflection of software package quality.
